---
title: "test"
description: |
  A short description of the post.
author:
  - name: Su Yiin Ang
    url: https://public.tableau.com/profile/suyiinang
date: 04-07-2021
output:
  distill::distill_article:
    self_contained: false
---



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = TRUE)
```

# 1. Introduction

## 1.1 Overview of application

The increasing availability of data has resulted in the increased demand for data driven decisions. Although there is an extensive range of commercial statistical tools, they are often subscription-based and demand good technical knowledge to mine and draw insights from. Therefore, it may not appeal to the average user.

Hence, the main idea of this application is its **user-friendly and interactive** R Shiny interface that would enable everyone to make data based decisions **without** needing programming or statistical background. We have selected 

There are three parts to this shiny application - Text Mining, Exploratory and Confirmatory, and Predictive Analytics. 

Please visit our website at https://ourshinypet.netlify.app/ for more details.

## 1.2 Objective of this report

This report will cover the **Exploratory and Confirmatory Analysis module**.

The main objective of this report is to  

(i) identify and select the appropriate R packages for the final R shiny application,  
(ii) develop prototypes of the module, and  
(iii) prepare storyboard sketch for the sub-module design.  

# 2. Literature review





# 3 Extracting, wrangling and preparing the input data 

Listings of Airbnbs in Singapore were extract from [InsideAirbnb.com](http://insideairbnb.com/get-the-data.html)

## 3.1 Load the relevant packages

Our focus is on using packages from the [`Tidyverse`](https://www.tidyverse.org/) family.

The following packages were mainly used in exploring and developing our prototype.  
- [`readr`](https://readr.tidyverse.org/), [`tibble`](https://tibble.tidyverse.org/), [`dplyr`](https://dplyr.tidyverse.org/), [`tidyr`](https://tidyr.tidyverse.org/) to load, process and prepare data for final exploration.   
- [`ggplot2`](https://ggplot2.tidyverse.org/index.html) to create exploratory plots.  
- [`plotly`](https://plotly.com/r/) to create interactive plots for exploratory analysis.   
- [`ggstatsplot`](https://indrajeetpatil.github.io/ggstatsplot/index.html) to create plots with statistical tests included within plot.  

```{r, message = FALSE, class.source = 'fold-show'}
packages <- c('tidyverse','ggstatsplot','plotly','readr','haven','funModeling','crosstalk','data.table','shiny', 'shinyWidgets', 'skimr', 'ggmosaic','ggExtra','ggpubr')

for (p in packages){
  if (!require(p,character.only=T)){
    install.packages(p)
  }
  library(p, character.only=T)
}
```

## 3.2 Extracting and wrangling the dataset

```{r,  message = FALSE, class.source = 'fold-show'}
# Read the csv file as a tbl_df
listing_prep <- read_csv("./data/listing_prep.csv")
listing_prep <- listing_prep %>%
  mutate(price, price_per_pax = round(price/accommodates,2)) %>%
  mutate_if(is.logical, as.character)

var_list <- sort(names(listing_prep))
listing_num <- select_if(listing_prep, is.numeric)
listing_cat <- listing_prep %>% select_if(~is.character(.) | is.logical(.))

set.seed(123) # for reproducibility

```

# 4. Testing protypes for submodule

## 4.1 Plotting univariate charts

For univariate charts, we would like to be able to observe :

i) Distribution - through histogram for numerical variables and barplot for categorical variables

ii) Outlier of selected variable - through boxplot 

### 4.1.1 Plotting distribution with histogram and bar plots

#### 4.1.1.1 With`ggstatsplot`](https://indrajeetpatil.github.io/ggstatsplot/index.html)

`ggbarstats()` currently only supports two variables as evidence in the [documentation](https://indrajeetpatil.github.io/ggstatsplot/reference/ggbarstats.html) of the `y` argument. 

See extract below:

> The variable to use as the columns in the contingency table. Please note that if there are empty factor levels in your variable, they will be dropped. Default is `NULL.` If `NULL`, one-sample proportion test (a goodness of fit test) will be run for the `x` variable. Otherwise an appropriate association test will be run. This argument can not be `NULL` for `ggbarstats` function.

Hence the visual below only apply to numerical variables.

```{r}
####################
#variable class
####################
output$d_type <- renderPrint({
  d_type <- unlist(listing_prep[,input$x_hist])
  paste('Datatype of selected variable' , class(d_type))})
####################
#gghistostats
####################

output$gghist <- renderPlot({
  
  x <- unlist(listing_prep[,input$x_hist2])
  
  hist <- gghistostats(
    data = listing_prep,
    x = !!colnames(listing_prep[input$x_hist]), 
    title = paste('Distribution of ', input$x_hist),
    normal.curve = TRUE,
    normal.curve.args = list(color = "red", size = 1),
    ggtheme = ggplot2::theme_classic()
    )
  hist
})

#################
#ui
###################
bscols(widths = c(12),
       list(
          selectInput(inputId = 'x_hist',
          label = 'Select x-variable',
          choices = sort(colnames(listing_num)),
          selected = 'price_per_pax'),
           
          verbatimTextOutput('d_type')),
        
          plotOutput('gghist')
 
)
```


#### Attempt to add interactivity to [`ggstatsplot`](https://indrajeetpatil.github.io/ggstatsplot/index.html) chart using [`ggplotly`](https://plotly.com/ggplot2/)

We have tried to wrap the `ggstatsplot` chart wwith `ggplotly()`, however by doing so, we would lose key metric such as the fitted normal curve and statistical test results. As clarified by the author [here](https://github.com/IndrajeetPatil/ggstatsplot/issues/403), this is a [`ggplot2`](https://ggplot2.tidyverse.org/reference/index.html) issue.

```{r}
renderPlotly({
  p <- gghistostats(
    listing_prep,
    x = !!colnames(listing_prep[input$x_hist]),
    xlab =  paste('Distribution of ', input$x_hist),
    title = 'ggstatsplot histogram and ggplotly wrapper for interactivity',
    normal.curve = TRUE,
    normal.curve.args = list(color = "red", size = 1),
    ggtheme = ggplot2::theme_classic())
  
  ggplotly(p)
  
})

```

***

#### 4.1.1.2 With [`ggplot2`](https://ggplot2.tidyverse.org/reference/index.html)

To work around `ggstatsplot` limitation, we have used `ggplot2` the visualise the variables and included computed statistical test below. 

For the statistical tests, we have assume normality as our dataset has 3000+ observations. Hence, we have used `t.test` for numerical variables and `prop.test` for categorical variables.

```{r}
###############
# themes
##############
themes <- list("Light" = theme_light(),
               "Minimal" = theme_minimal(),
               "Classic" = theme_classic(),
               "B&W" = theme_bw(),
               "Gray" = theme_gray())
plot_theme <- reactive({themes[[input$iTheme]]})
####################
#variable class
####################
output$d_type2 <- renderPrint({
  d_type <- unlist(listing_prep[,input$x_hist2])
  paste(class(d_type))})
#####################
# histogram / barplots
######################

 output$dist <- renderPlotly({
   x <- unlist(listing_prep[,input$x_hist2])
   
   base <- ggplot(listing_prep, aes(x = x)) + 
     labs(title = paste("Distribution of ", input$x_hist2)) +
     xlab(paste(input$x_hist2)) +
     plot_theme()
   
   num <- geom_histogram(bins = input$bin,
                          color = '#767676',
                          fill = '#FF5A5F')
      
   cat <- geom_histogram(stat = 'count',
                       color = '#767676',
                          fill = '#FF5A5F')
   
   hist <- if(class(x) == "numeric"){
     base + num 
   } else {
     base + cat
   }

   hist_mean <- geom_vline(aes(xintercept=mean(x,na.rm=T)),color="#00A699", linetype="dashed", size=1)
   hist_median <- geom_vline(aes(xintercept=median(x,na.rm=T)),color="#484848", linetype="dashed", size=1)

   show_m <- hist + hist_mean + hist_median

  ggplotly(show_m)   
  
  })
#####################
#conf test
####################
output$conf_test <- renderPrint({
  
  x <- unlist(listing_prep[,input$x_hist2])
  
  num <- t.test(x,
         alternative = input$alternative,
         conf.level = input$conf_lev_hist)

  test <-if(class(x) == 'numeric'){num}
  else if (class(x) == 'logical'){return('One proportion z-test is under work in progress')}
  else if (class(x) == 'character'){return('One proportion z-test is under work in progress')}
  else {return('One proportion z-test is under work in progress')}
  
  test
})
###################
# uii
#################
bscols(widths = c(3,9),
       list(selectInput(inputId = 'x_hist2',
          label = 'Select x-variable',
          choices = var_list,
          selected = 'price_per_pax'),
          
        verbatimTextOutput('d_type2'),

        sliderInput(inputId = "bin", 
                          label = "Number of bins", 
                          min = 5, 
                          max = 30, 
                          value = 10),
        
        selectInput(inputId = 'iTheme',
                             label = 'Select theme',
                             choices = sort(names(themes)),
                             selected = 'Gray'),
        
       selectInput(inputId = 'alternative',
                       label = 'Alternative',
                       choices = c('two.sided','less','greater'),
                       selected = 'two.sided'),
       
        sliderInput(inputId = 'conf_lev_hist',
            label = "Confidence Interval",
            min = 0,
            max = 1,
            value = 0.95)

       ),
       plotlyOutput('dist'))

verbatimTextOutput('conf_test')
 
```


### 4.1.2 Plotting outliers with boxplot

[`ggstatsplot`](https://indrajeetpatil.github.io/ggstatsplot/index.html) currently doesn't not have any graph to plot a single variable boxplot for outlier observations. As such, I have used [`ggplotly()`](https://plotly.com/ggplot2/) and [`ggplot2`](https://ggplot2.tidyverse.org/reference/index.html)  to create and interactive chart. 

***

#### 4.2.1 Using `ggplot2`

```{r}
bscols(widths = c(12),
       list(
        strong('Summary statistics for review_scores_rating'),
        
        renderPrint({summary(listing_prep$review_scores_rating)}),
       
        renderPlotly({
          boxchart <- ggplot(listing_prep, 
                             aes(x = '', y = review_scores_rating, colour = '#FF5A5F')) +
            geom_boxplot() +
            coord_flip() +
            stat_boxplot(geom ='errorbar') +
            stat_summary(fun.y=mean, geom="point", shape=5, size=4)+ 
            labs(title = paste("Outlier boxplot using ggplot and ggplotly")) +
            xlab('review_scores_rating') +
            theme_classic() +
            theme(legend.position = 'none')
            
            boxchart
          })
))
         
```

## 4.2 Bivariate exploratory and confirmatory analysis

For bivariate analysis, we would like to be able to plot 3 main types of interactions between variables - 

- 2 numerical variables - Scatterplot 

- 1 numerical and 1 categorical - box/violin plot

- 2 categorical variables - mosaic plot

*** 

### 4.2.1 Plotting numerical variables with scatterplot

#### 4.2.1.1 using `ggMarginal`

- add marginal plots to [`ggplot2`](https://ggplot2.tidyverse.org/reference/index.html) charts but not interactive - doesn't not interact with plotly.

```{r}
renderPlot({
  p1 <- ggplot(listing_prep, 
               aes(host_listings_count, review_scores_rating, colour = host_is_superhost)) +
  geom_point() +
  title('Scatterplot of host_listings_count and review_scores_rating using ggMarginal')

  ggMarginal(p1, groupColour = TRUE, groupFill = TRUE)
})
```

***

#### 4.2.1.2 using [`ggstatsplot`](https://indrajeetpatil.github.io/ggstatsplot/index.html)

```{r}


bscols(widths = c(3,9),
       list(
          selectInput(inputId = 'x_scat',
          label = 'Select x-variable',
          choices = colnames(listing_num),
          selected = 'host_listings_count'),
          
          selectInput(inputId = 'y_scat',
          label = 'Select y-variable',
          choices = colnames(listing_num),
          selected = 'review_scores_rating'),
          
          sliderInput(inputId = 'conf_lev',
            label = "Confidence Interval",
            min = 0,
            max = 1,
            value = 0.95)),
        
          renderPlot({
          ggscatterstats(
            data = listing_prep,
            x = !!colnames(listing_prep[input$x_scat]), 
            y = !!colnames(listing_prep[input$y_scat]), 
            conf.level = input$conf_lev,  
            title = paste('Scatterplot of ', input$x_scat, ' and ', input$y_scat, 'using ggstatsplot')
            )
          })
  
)

```

***

#### 4.2.1.3 using [`ggplot2`](https://ggplot2.tidyverse.org/reference/index.html) and [`ggplotly()`](https://plotly.com/ggplot2/)

```{r}

output$correlation <- renderPrint({
  x_num <- unlist(listing_prep[,input$x_num])
  
  y_num <- unlist(listing_prep[,input$y_num])
  
  print_cor <- cor.test(x_num, y_num)
  
  print_cor
  
})

output$scatter <- renderPlotly({
  x_num <- unlist(listing_prep[,input$x_num])
  
  y_num <- unlist(listing_prep[,input$y_num])
  
  c_cat <- if(input$scatter_colour == 'None'){
    'None'} else { unlist(listing_prep[,input$scatter_colour]) }
  
  scatter <- ggplot(listing_prep, aes(x = x_num, y = y_num)) + 
    labs(
      title = paste("Scatterplot of ", input$x_num, 'and', input$y_num),
      x = paste(input$x_num),
      y = paste(input$y_num),
      colour = paste(input$scatter_colour)) 
  
  add_colour <- geom_point(aes(fill = c_cat), alpha = 0.8)
  
  show_colour <- if(input$scatter_colour == 'None' ){
     scatter + geom_point()
   } else {
     scatter + add_colour
   }
  
  show_smooth <- if(input$smoothline){
     show_colour + geom_smooth(method = "lm", se = FALSE)
   } else {
     show_colour
   }  
  
  scatterly <- ggplotly(show_smooth)
  
  scatterly
})

bscols(widths = c(3,9),
       list(
        
        selectInput(inputId = 'x_num',
          label = 'Select x-variable',
          choices = sort(colnames(listing_num)),
          selected = 'review_scores_rating'),
        
        selectInput(inputId = 'y_num',
          label = 'Select y-variable',
          choices = sort(colnames(listing_num)),
          selected = 'amenities_count'),
        
        selectInput(inputId = 'scatter_colour',
                   label = 'Colour chart by:',
                   choices = c('None', sort(colnames(listing_cat))),
                   selected = 'None'),
        
        checkboxInput(inputId = 'smoothline',
                      label = 'Add best fit line',
                      value = TRUE)
       ),
       plotlyOutput('scatter')
       )

verbatimTextOutput('correlation')
```

***

### 4.2.2 Plotting numerical and categorical variabes - Box and violin plot

#### 4.2.2.1 Using [`ggstatsplot`](https://indrajeetpatil.github.io/ggstatsplot/index.html)

```{r}

bscols(widths = c(3,9),
  list(selectInput(inputId = 'x_cda',
                   label = 'Select x-variable',
                   choices = sort(colnames(listing_cat)),
                   selected = 'host_is_superhost'),
       
       selectInput(inputId = 'y_cda',
                   label = 'Select y-variable',
                   choices = sort(colnames(listing_num)),
                   selected = 'price_per_pax'),
       
       selectInput(inputId = 'type',
                   label = 'Select statistical approach ',
                   choices = c("parametric", 'nonparametric','robust','bayes'),
                   selected = 'parametric'),
       
       selectInput(inputId = 'pairwise',
                   label = 'Select pairwise option ',
                   choices = c("significant", 'non-significant','all'),
                   selected = 'significant'),
       
       selectInput(inputId = 'p_adj',
                   label = 'Select adjustment method of p-values ',
                   choices = c("holm", 'hochberg','hommel','bonferroni', 'BH', 'BY', 'fdr', 'none'),
                   selected = 'holm'),
       
       sliderInput(inputId = 'conf_lev2',
                   label = "Confidence Interval",
                   min = 0,
                   max = 1,
                   value = 0.95)),
       
      renderPlot({ggbetweenstats(
        data = listing_prep,
        x = !!colnames(listing_prep[input$x_cda]), 
        y = !!colnames(listing_prep[input$y_cda]), 
        title = paste('Comparison of ', input$y_cda, ' by ', input$x_cda),
        type = input$type,
        conf.level = input$conf_lev2,
        pairwise.comparisons = TRUE,
        pairwise.display = input$pairwise, #not working
        p.adjust.method = input$p_adj, #not working
        ggtheme = ggplot2::theme_classic())
        })
  
)
```

***

#### 4.2.2.2 Using [`ggplot2`](https://ggplot2.tidyverse.org/reference/index.html) and [`ggplotly()`](https://plotly.com/ggplot2/)

```{r}
output$bbox <- renderPlotly({
  x_b_cat <- unlist(listing_prep[,input$x_b_cat])
  
  y_num <- unlist(listing_prep[,input$y_b_num])
  
  ccat <- if(input$box_colour == 'None'){'None'} 
          else {unlist(listing_prep[,input$box_colour]) }

  base <- ggplot(listing_prep, aes(x = x_b_cat, y = y_num)) + 
    labs(
      title = paste(input$chart_type, 'plot of ', input$x_b_cat, ' and ', input$y_b_num),
      x = paste(input$x_b_cat),
      y = paste(input$y_b_num),
      colour = paste(input$box_colour))
  
  add_box_c <- geom_boxplot(aes(fill = ccat), outlier.shape = NA)
  add_box <- geom_boxplot(outlier.shape = NA)
  add_vio <- geom_violin()
  add_vio_c <- geom_violin(aes(fill = ccat))
  
  bbox <- if (input$box_colour != 'None' & input$chart_type == 'Box'){base + add_box_c}
          else if(input$box_colour == 'None' & input$chart_type == 'Box'){ base + add_box }
          else if(input$box_colour != 'None' & input$chart_type == 'Violin'){ base + add_vio_c }
          else { base + add_vio}
  
  
  flip_chart <- if(input$flipxy){
     bbox + coord_flip()
   } else {
     bbox
   }  
  
  boxly <- ggplotly(flip_chart) %>% layout(boxmode = "group")
  
  boxly
})


bscols(widths = c(3,9),
       list(
        
        selectInput(inputId = 'x_b_cat',
          label = 'Select x-variable',
          choices = sort(colnames(listing_cat)),
          selected = 'review_scores_rating'),
        
        selectInput(inputId = 'y_b_num',
          label = 'Select y-variable',
          choices = sort(colnames(listing_num)),
          selected = 'price_per_pax'),
        
        selectInput(inputId = 'box_colour',
                   label = 'Colour chart by:',
                   choices = c('None',sort(colnames(listing_cat))),
                   selected = 'host_is_superhost'),
        
        selectInput(inputId = 'chart_type',
                    label = 'Select chart type:',
                    choices = c('Box', 'Violin'),
                    selected = 'box'),
        
        checkboxInput(inputId = 'flipxy',
              label = 'Flip axis',
              value = FALSE)
        
       ),
       plotlyOutput('bbox'))
```

***

### 4.3.3. Plotting categorical variables - Mosaic plot




#### 4.3.3.1 Using [`ggstatsplot`](https://indrajeetpatil.github.io/ggstatsplot/index.html)

```{r}
bscols(widths = c(12),
       list(
        
        selectInput(inputId = 'x_m_cat1',
          label = 'Select x-variable',
          choices = sort(colnames(listing_cat)),
          selected = 'host_is_superhost'),
        
        selectInput(inputId = 'y_m_cat1',
          label = 'Select y-variable',
          choices = sort(colnames(listing_cat)),
          selected = 'property_type'),
       
         sliderInput(inputId = 'conf_lev_bar',
            label = "Confidence Interval",
            min = 0,
            max = 1,
            value = 0.95),
        
        selectInput(inputId = 'propt',
          label = 'Proportion test',
          choices = c('Yes' = TRUE,
                      'No' = FALSE),
          selected = 'Yes')),
        
        renderPlot({
           ggbarstats(
              data = listing_prep,
              x = !!colnames(listing_prep[input$x_m_cat1]), 
              y = !!colnames(listing_prep[input$y_m_cat1]), 
              title = paste('Mosaic plot using ggstatsplot'),
              type = input$type,
              conf.level = input$conf_lev_bar,
              proportion.test = input$propt,
              ggtheme = ggplot2::theme_classic()
              )})
       )
```

***

#### 4.3.3.2 Using [`ggmosaic`](https://haleyjeppson.github.io/ggmosaic/), [`ggplot2`](https://ggplot2.tidyverse.org/reference/index.html) and [`ggplotly()`](https://plotly.com/ggplot2/)

```{r}
output$mosaic <- renderPlotly({
  
  mosaic_c <- ggplot(listing_prep) + 
  geom_mosaic(aes(x=product(!!sym(input$x_m_cat), !!sym(input$y_m_cat)), fill = !!sym(input$x_m_cat))) + 
    labs(
      title = paste("Mosaic plot using ggmosaic and plotly"),
      x = paste(input$x_m_cat),
      y = paste(input$y_m_cat)) + 
    theme(axis.text.y=element_blank(),
        axis.ticks.y=element_blank(),
        axis.text.x = element_text(angle = 90))
  
  ggplotly(mosaic_c)
  
})

output$chisqtest <- renderPrint({
  
  x_var <- unlist(listing_prep[,input$x_m_cat])
  y_var <- unlist(listing_prep[,input$y_m_cat])
  
  chisq <- chisq.test(x = x_var, y = y_var)
  chisq
  
})

bscols(widths = c(12),
       list(
        
        selectInput(inputId = 'x_m_cat',
          label = 'Select x-variable',
          choices = sort(colnames(listing_cat)),
          selected = 'host_is_superhost'),
        
        selectInput(inputId = 'y_m_cat',
          label = 'Select y-variable',
          choices = sort(colnames(listing_cat)),
          selected = 'property_type'),
        
       plotlyOutput('mosaic'),
       strong('Association test'),
       verbatimTextOutput('chisqtest')
       ))
```

# Outcome - final storyboard for Exploratory

`ggstatsplot` is an amazing package, especially for those without extensive statistical background. However, given that it is a relatively new package, there are some outstanding compatability issues with Shiny as raised in [Github](https://github.com/IndrajeetPatil/ggstatsplot/issues/475). We experienced some issues when attempting to sync the shinywidgets to the chart.

Although `ggplot2` lacks the automation of statistical insights, it has an extensive list of charts available, interactive with `ggplotly()` and, more importantly, compatible with R Shiny. The lack of statistical automation can be mitigated through performing the test at the server backend. 

Given the above issue with `ggstatsplot` and the need for interactivity, our final R Shiny application will use charts from the `ggplot2` package and statistical tests will be automated at the backend.



